Boundary terms in the action for Causal Sets

### 2014-10-14 paper_v7.tex
more changes

Consider a set of synchronous or Gaussian Normal Coordinates (GNC), $x^\mu= (t,\mathbf x)$, adapted to $\Sigma$ such that in neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
\be
ds^2 = -dt^2 + h_{ij} (t,\mathbf x) dx^i dx^j.
\ee
In these coordinates, the surface $\Sigma$ corresponds to $t=0$, and the coordinate $t$ measures the proper time elapsed along geodesics whose tangent vector is proportional to the surface normal on $\Sigma$.
%\be\label{GNC_metric}
%g_{\mu\nu} (x)=
%\begin{pmatrix}
 %-1&0 \\
% 0&h_{ij} (x)
%\end{pmatrix}.
%\ee
%In order to find $N_{max}$, the number of maximal points below the surface, one has to use the fact that the points have been sprinkled with a Poisson distribution. 
%If we pick a point $x_0= (0,\mathbf{x})\in \Sigma$ on the surface that has the same spatial coordinates as $x= (t,\mathbf{x})$ then the coordinate time, $t$, is the proper time elapsed along the unique geodesic between $x_0$ and $x$. The volume $V (\Sigma,x)$ then only depends on $t$.~\mb{error?}  
The integrals~\eqref{eq:nmax} and~\eqref{eq:nmin} seem intractable as they stand, since the integration is over the entire causal past/future of the surface and the volume expressions will be complicated in the presence of curvature. However, due to $\Sigma$ being closed compact and $M^+$ and $M^-$ of finite volume, we can always find a subneighbourhood of $U_\Sigma$ such that the contribution to the boundary term from points outside of that neighbourhood tends to zero exponentially quickly as $\rho\rightarrow\infty$.\footnote{To see this, let us restrict ourselves to points to the future of $\Sigma$ and hence to the contributions to $\mathbf F_k$. In the GNC system, there will be some $t_{max}>0$ such that all points $(t,\mathbf x)$ with $|t|<t_{max}$ lie in $U_\Sigma$. Choose some $t'<t_{max}$, and consider the subneighbourhood $U'_\Sigma=\left\{(t,\mathbf x)\in U_\Sigma: |t|<t'\right\}$. Then given any $\Lambda\in\mathbb R^+$, there exists, for each $\mathbf x$, a value $\rho_\Lambda(\mathbf x)$ such that $V_\blacktriangle(\mathbf x, t)\rho_\Lambda(\mathbf x)>\Lambda$ for all points outside of $U'_\Sigma$. Since $\Sigma$ is compact, $\rho_\Lambda(\mathbf x)$ achieves its maximum on $\Sigma$ and we can choose a finite constant value $\rho_\Lambda\equiv\max_{\mathbf{x}}\rho_\Lambda(\mathbf x)$ such that $\rho_\Lambda V(\mathbf x, t)>\Lambda$ for all $x\in M^+\setminus U'_\Sigma$. Since $M^+$ has finite volume, the total contribution from those points to $\left<\mathbf P_k\right>$ is then bounded from above by $\rho_\Lambda V(M^+\setminus U'_\Sigma)\frac{\Lambda^k}{k!}e^{-\Lambda}$ for large enough $\Lambda$, a bound that can be made arbitrarily small.} 
Hence, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero. These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but so long as $\Sigma$ is a closed compact manifold and is everywhere spacelike then all we require is that the radius of extrinsic curvature, $R$, satisfies $R\gg l$. The integrals then simplify to
\begin{gather}\label{eq:nmax_and_eq:nmin}
\begin{aligned}
\left\langle \BF{k}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
h^{\frac{1}{2}}\left (1+
\frac{1}{2}\frac{\dot{h}}{h}t+O (t^2)\right)
 \frac{\left (\rho\: V_\blacktriangledown (p)\right)^k}{k!} e^{-\rho V_\blacktriangledown (t,\mathbf x)}
\\
\left\langle \BP{k}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left (1+
\frac{1}{2}\frac{\dot{h}}{h}t+O (t^2)\right) \frac{\left (\rho\: V_\blacktriangle (p)\right)^k}{k!} e^{-\rho V_\blacktriangle (t,\mathbf x)}
\end{aligned}
\end{gather}
where $h:= det\left (h_{ij} (0,\mathbf{x})\right)$ and $\dot{}\: := \frac{\partial}{\partial t}$ and the metric determinant has been in expanded in small $t$.

@@

Fix $p\in M$ and denote its coordinate values in GNCs by $x^\mu_p= (t_p,\mathbf x_p)$. It will be convenient to use RNCs centered not at the tip, $p$, of the cone but instead at the point $p_0$ where the unique geodesic through $p$ whose tangent is normal to $\Sigma$ intersects $\Sigma$. In GNCs this simply corresponds to the point $p_0$ with coordinates $x_0^\mu=x^\mu (p_0)= (0,\mathbf x_p)$. RNCs centered at $p_0$ will be given the symbol $y^{\overline{\mu}}$, or in terms of coordinates as $(y^{\overline{0}},\mathbf{y})=(\overline{t},\mathbf{y})$.
%\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} 
We need to assume that for any $p$ in $U_\varepsilon$, the Riemann normal neighbourhood $U_p\subset M$ (throughout which the RNC system centered at $p_0$ is well-defined) contains the truncated cone $J^- (p)\cap J^+ (\Sigma)$. This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$. In RNCs the metric and the Christoffel symbols at $p_0$ are those of flat space:
\be\label{eq:RNCMetricTransAtPAndChris}
g_{\overline{\mu} \overline{\nu}} (p_0)=\eta_{\overline{\mu} \overline{\nu}}=A^{\mu}_{\;\overline{\mu}}A^{\nu}_{\;\overline{\nu}}g_{\mu\nu} (p_0)\;,\;\;\;\;\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0
\ee
The $A^{\mu}_{\;\overline{\mu}}$ govern the coordinate transformation from GNCs to RNCs to linear order. %but $O (x^2)$ corrections may be required. 
To second order the coordinate transformation is given by
\be\label{eq:RNCtotaltrans}
y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}x^\nu+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)x^\nu x^\rho+O ( (x-x_0)^3).
\ee
The inverse relation to first order is 
\be\label{eq:RNCinversetrans}
x^{\mu}=A^{\mu}_{\;\overline{\nu}}y^{\overline{\nu}}+O (y^2)
\ee
and one finds that the $A^{\mu}_{\;\overline{\mu}}$ must satisfy
\be\label{eq:RNCeqnforA}
A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}\;,\;\;\;\;A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}
\ee
Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$. For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=t_p=T$.

@@

This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}_p$ to write down the integration limits. There are two parts to the boundary of the (solid) truncated cone: the \emph{base}, defined as $B(p):=\Sigma\cap J^-(p)$, and the \emph{top} defined as the set corresponding to the lightcone emanating at $p$ with all points below $\Sigma$ removed, written as $T(p):=\partial J^-(p)\setminus\partial(J^-(p)\cap (J^-(\Sigma)\setminus\Sigma))$. Now the boundary of $\mathcal{X}_p$ is simply $\partial\mathcal{X}_p=B(p)\cup T(p)$. First we look at the top. Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone volume comes in at $O (T^{d+2})$ and so can be ignored for our purposes. This means that the top can be treated as an effectively flat lightcone in RNCs and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.

The base of the cone in GNCs is simply a part of the surface $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNCs. Equation (\ref{eq:RNCtotaltrans}) gives
\be\label{eq:BottomSurfaceWithGNC}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)x^i x^j+O ( x^3)
\ee
The linear part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface. Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:
\be\label{eq:BottomSurface}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}y^{\ibar} y^{\jbar}+O (y^3)
\ee
Let us rewrite this equation in spherically symmetric coordinates, i.e. define $r:=\sqrt{\delta_{\ibar\jbar}y^\ibar y^\jbar}$ and the usual angular coordinates $\phi_1,..,\phi_{d-2}$ in terms of the spatial coordinates $y^{\overline{1}} = r \cos (\phi_1),\ldots, y^{\overline{d-1}} = r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})$. Then
\be\label{eq:RadialBottomSurface}
\overline{t}=\frac{1}{2}\left (\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\frac{y^{\ibar} y^{\jbar}}{r^2}\right)r^2+O (y^3)=\frac{1}{2}f (\mathbf{x}_p,{\boldsymbol\phi})r^2+O (y^3)
\ee
where $\boldsymbol\phi$ stands collectively for all the angular coordinates $\phi_1,..,\phi_{d-2}$. The function $f (\mathbf{x}_p,\boldsymbol\phi)$ depends on $\mathbf{x}_p$ since $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\ibar}$ depend on $p_0$.
%One can see that $f (\mathbf{x}_0,\boldsymbol\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})
%\begin{equation}
%\begin{aligned}\label{eq:SphericalCoords}
%y^{\overline{1}} &= r \cos (\phi_1)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \cos (\phi_2)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \sin (\phi_2) \cos (\phi_3)  \\
%    &\vdots  \\
%y^{\overline{d-2}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \cos (\phi_{d-2})  \\
%y^{\overline{d-1}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})
%\end{aligned}
%\end{equation}

With the boundaries of the integration region in place, we can now write down the integral explicitly in spherical coordinates:
\be\label{eq:VolumeIntegralSpherical}
V_\blacktriangle (p)=\int_{S^{d-2}}
d\Omega_{d-2}
\int_{0}^{r_{max} (\phi)}r^{d-2}dr
\int_{\frac{1}{2}f (\mathbf{x}_p,\phi)r^2}^{-r+T}
d\overline{t}+O (T^{d+2})
\ee
where $r_{max} (\boldsymbol\phi)$ is the value of the radial coordinate for which the base intersects the top at an angle $\boldsymbol\phi$, as shown in Figure \ref{fig:cone_plot}. To find this value we equate the time coordinates of the top and base surfaces to give 
\be
\frac{1}{2}f (\mathbf{x}_p,\boldsymbol\phi){r_{max}}^2 (\boldsymbol\phi)=-r_{max} (\boldsymbol\phi)+T
\ee 
We then solve this for $r_{max} (\boldsymbol\phi)$ and take the positive solution. The solution can be expanded in $T$ and is simply $r_{max}=T+O (T^2)$, with angular dependent terms contributing at $O (T^2)$. The $O (T^2)$ term will contribute at $O (T^{d+2})$ in the volume integral and so can be ignored. Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) allows us to evaluate the integral, wherein we find 
\begin{figure}[t]
  \centering
    {\includegraphics[scale=0.5]{coneplot2}}
     \caption{A $3$-dimensional representation of the region $\mathcal{X}_p$ in RNCs. The top part of $\partial\mathcal{X}_p$ can be approximated as a flat cone, and the base surface intersects the top part at a radial coordinate, $r_{max} (\boldsymbol\phi)$, which will in general be a function of the angles $\boldsymbol\phi$ (in $3$-dimensions there is one angle $\phi$). This function is found by projecting down from the intersection to the $\overline{t}$ plane. The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
     \label{fig:cone_plot}
\end{figure}
\be\label{eq:VolumeNoK}
V_\blacktriangle (p)
=\frac{V_{d-1}}{d}T^d\left (1-\frac{d}{2 (d+1)}\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}T\right)
+O (T^{d+2})
\ee
where the $\delta^{\ibar\jbar}$ comes from the fact that cross terms ($\ibar\neq \jbar$) vanish under the angular integration. The defining relations for $A^{i}_{\;\ibar}$ can now be rearranged to give $A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}=h^{ij} (p_0)$. In GNCs the extrinsic curvature on the surface is given by
\be\label{eq:K}
K
=g^{\mu\nu }\nabla_{\mu}n_{\nu}
=-\Gamma^{0}_{\;ij}h^{ij}=-\frac{1}{2}\frac{\dot{h}}{h}.
\ee
Substituting this into~\eqref{eq:VolumeNoK} we obtain 
\begin{align}
V_\blacktriangle (T,\mathbf x)
&=\frac{V_{d-1}}{d}T^d\left (1+\frac{d}{2 (d+1)}K (0,\mathbf{x})T\right)
+O (T^{d+2}) \label{eq:TopVolumeWithK}\\
V_\blacktriangledown (-T,\mathbf x)
&=\frac{V_{d-1}}{d}T^d\left (1-\frac{d}{2 (d+1)}K (0,\mathbf{x})T\right)
+O (T^{d+2}), \label{eq:BottomVolumeWithK}
\end{align}
having dropped the subscript $p$. Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\left\langle \BP{k}\right\rangle$ and $\left\langle \BF{k}\right\rangle$.

>>>>

Consider a set of synchronous or Gaussian Normal Coordinates (GNC), $x^\mu= (t,\mathbf x)$, adapted to $\Sigma$ such that in neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
\be
ds^2 = -dt^2 + h_{ij} (t,\mathbf x) dx^i dx^j.
\ee
In these coordinates, the surface $\Sigma$ corresponds to $t=0$, and the coordinate $t$ measures the proper time elapsed along geodesics whose tangent vector is proportional to the surface normal on $\Sigma$.
%\be\label{GNC_metric}
%g_{\mu\nu} (x)=
%\begin{pmatrix}
 %-1&0 \\
% 0&h_{ij} (x)
%\end{pmatrix}.
%\ee
%In order to find $N_{max}$, the number of maximal points below the surface, one has to use the fact that the points have been sprinkled with a Poisson distribution. 
%If we pick a point $x_0= (0,\mathbf{x})\in \Sigma$ on the surface that has the same spatial coordinates as $x= (t,\mathbf{x})$ then the coordinate time, $t$, is the proper time elapsed along the unique geodesic between $x_0$ and $x$. The volume $V (\Sigma,x)$ then only depends on $t$.~\mb{error?}  
The integrals~\eqref{eq:nmax} and~\eqref{eq:nmin} seem intractable as they stand, since the integration is over the entire causal past/future of the surface and the volume expressions will be complicated in the presence of curvature. However, due to $\Sigma$ being closed compact and $M^+$ and $M^-$ of finite volume, we can always find a subneighbourhood of $U_\Sigma$ such that the contribution to the boundary term from points outside of that subneighbourhood tends to zero exponentially quickly as $\rho\rightarrow\infty$.\footnote{To see this, let us restrict ourselves to points to the future of $\Sigma$ and hence to the contributions to $\mathbf F_k$. In the GNC system, there will be some $t_{max}>0$ such that all points $(t,\mathbf x)$ with $|t|<t_{max}$ lie in $U_\Sigma$. Choose some $t'<t_{max}$, and consider the subneighbourhood $U'_\Sigma=\left\{(t,\mathbf x)\in U_\Sigma: |t|<t'\right\}$. Then given any $\Lambda\in\mathbb R^+$, there exists, for each $\mathbf x$, a value $\rho_\Lambda(\mathbf x)$ such that $V_\blacktriangle(\mathbf x, t)\rho_\Lambda(\mathbf x)>\Lambda$ for all points outside of $U'_\Sigma$. Since $\Sigma$ is compact, $\rho_\Lambda(\mathbf x)$ achieves its maximum on $\Sigma$ and we can choose a finite constant value $\rho_\Lambda\equiv\max_{\mathbf{x}}\rho_\Lambda(\mathbf x)$ such that $\rho_\Lambda V(\mathbf x, t)>\Lambda$ for all $x\in M^+\setminus U'_\Sigma$. Since $M^+$ has finite volume, the total contribution from those points to $\left<\mathbf F_k\right>$ is then bounded from above by $\rho_\Lambda V(M^+\setminus U'_\Sigma)\frac{\Lambda^k}{k!}e^{-\Lambda}$ for large enough $\Lambda$, a bound that can be made arbitrarily small.} 
Hence, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero. These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but so long as $\Sigma$ is a closed compact manifold and is everywhere spacelike then all we require is that the radius of extrinsic curvature, $R$, satisfies $R\gg l$. The integrals then simplify to
\begin{gather}\label{eq:nmax_and_eq:nmin}
\begin{aligned}
\left\langle \BF{k}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
h^{\frac{1}{2}}\left (1+
\frac{1}{2}\frac{\dot{h}}{h}t+O (t^2)\right)
 \frac{\left (\rho\: V_\blacktriangledown (p)\right)^k}{k!} e^{-\rho V_\blacktriangledown (t,\mathbf x)}
\\
\left\langle \BP{k}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left (1+
\frac{1}{2}\frac{\dot{h}}{h}t+O (t^2)\right) \frac{\left (\rho\: V_\blacktriangle (p)\right)^k}{k!} e^{-\rho V_\blacktriangle (t,\mathbf x)}
\end{aligned}
\end{gather}
where $h:= det\left (h_{ij} (0,\mathbf{x})\right)$ and $\dot{}\: := \frac{\partial}{\partial t}$ and the metric determinant has been in expanded in small $t$.

@@

Fix $p\in M$ and denote its coordinate values in GNCs by $x^\mu_p= (t_p,\mathbf x_p)$. It will be convenient to use RNCs centered not at the tip, $p$, of the cone but instead at the point $p_0$ where the unique geodesic through $p$ whose tangent is normal to $\Sigma$ intersects $\Sigma$. In GNCs this simply corresponds to the point $p_0$ with coordinates $x_0^\mu=x^\mu (p_0)= (0,\mathbf x_p)$. RNCs centered at $p_0$ will be given the symbol $y^{\overline{\mu}}$, or in terms of coordinates as $(y^{\overline{0}},\mathbf{y})=(\overline{t},\mathbf{y})$.
%\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} 
We need to assume that for any $p$ in $U_\varepsilon$, the Riemann normal neighbourhood $U_p\subset M$ (throughout which the RNC system centered at $p_0$ is well-defined) contains the truncated cone $J^- (p)\cap J^+ (\Sigma)$. This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$. In RNCs the metric and the Christoffel symbols at $p_0$ are those of flat space:
\be\label{eq:RNCMetricTransAtPAndChris}
g_{\overline{\mu} \overline{\nu}} (p_0)=\eta_{\overline{\mu} \overline{\nu}}=A^{\mu}_{\;\overline{\mu}}A^{\nu}_{\;\overline{\nu}}g_{\mu\nu} (p_0)\;,\;\;\;\;\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0
\ee
The $A^{\mu}_{\;\overline{\mu}}$ govern the coordinate transformation from GNCs to RNCs to linear order. %but $O (x^2)$ corrections may be required. 
To second order the coordinate transformation is given by
\be\label{eq:RNCtotaltrans}
y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}(x^\nu -x_0^{\nu})+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)(x^\nu -x_0^{\nu}) (x^\rho -x_0^{\rho})+O ( (x-x_0)^3).
\ee
Where we can see that when $x^{\mu}=x_0^{\mu}$ we have that $y^{\overline{\mu}}=0$, and hence that our coordinates are centred at $p_0$. The inverse relation to first order is
\be\label{eq:RNCinversetrans}
x^{\mu}=x_0^{\mu}+A^{\mu}_{\;\overline{\nu}}y^{\overline{\nu}}+O (y^3)
\ee
where the $O(y^2)$ term vanishes due to the fact that $\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0$. One finds that the $A^{\mu}_{\;\overline{\mu}}$ must satisfy
\be\label{eq:RNCeqnforA}
A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}\;,\;\;\;\;A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}
\ee
Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$. For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$.

@@

This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}_p$ to write down the integration limits. There are two parts to the boundary of the (solid) truncated cone: the \emph{base}, defined as $B(p):=\Sigma\cap J^-(p)$, and the \emph{top} defined as the set corresponding to the lightcone emanating at $p$ with all points below $\Sigma$ removed, written as $T(p):=\partial J^-(p)\cap J^+(\Sigma)$. Now the boundary of $\mathcal{X}_p$ is simply $\partial\mathcal{X}_p=B(p)\cup T(p)$. First we look at the top. Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone volume comes in at $O (T^{d+2})$ and so can be ignored for our purposes. This means that the top can be treated as an effectively flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.

The base of the cone in GNCs is simply a part of the surface $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNCs. Equation (\ref{eq:RNCtotaltrans}) gives
\be\label{eq:BottomSurfaceWithGNC}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)(x^i-x_0^i) (x^j-x_0^j)+O ( (x-x_0)^3)
\ee
The linear part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}(x^{\mu}-x_0^{\mu})=t$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $t=0$ for the bottom surface. Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:
\be\label{eq:BottomSurface}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}y^{\ibar} y^{\jbar}+O (y^3)
\ee
Let us rewrite this equation in spherically symmetric coordinates, i.e. define $r:=\sqrt{\delta_{\ibar\jbar}y^\ibar y^\jbar}$ and the usual angular coordinates $\phi_1,..,\phi_{d-2}$ in terms of the spatial coordinates $y^{\overline{1}} = r \cos (\phi_1),\ldots, y^{\overline{d-1}} = r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})$. Then
\be\label{eq:RadialBottomSurface}
\overline{t}=\frac{1}{2}\left (\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\frac{y^{\ibar} y^{\jbar}}{r^2}\right)r^2+O (y^3)=\frac{1}{2}f (\mathbf{x}_p,{\boldsymbol\phi})r^2+O (y^3)
\ee
where $\boldsymbol\phi$ stands collectively for all the angular coordinates $\phi_1,..,\phi_{d-2}$. The function $f (\mathbf{x}_p,\boldsymbol\phi)$ depends on $\mathbf{x}_p$ since $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\ibar}$ depend on $p_0$.
%One can see that $f (\mathbf{x}_0,\boldsymbol\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})
%\begin{equation}
%\begin{aligned}\label{eq:SphericalCoords}
%y^{\overline{1}} &= r \cos (\phi_1)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \cos (\phi_2)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \sin (\phi_2) \cos (\phi_3)  \\
%    &\vdots  \\
%y^{\overline{d-2}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \cos (\phi_{d-2})  \\
%y^{\overline{d-1}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})
%\end{aligned}
%\end{equation}

With the boundaries of the integration region in place, we can now write down the integral explicitly in spherical coordinates:
\be\label{eq:VolumeIntegralSpherical}
V_\blacktriangle (p)=\int_{S^{d-2}}
d\Omega_{d-2}
\int_{0}^{r_{max} (\phi)}r^{d-2}dr
\int_{\frac{1}{2}f (\mathbf{x}_p,\phi)r^2}^{-r+T}
d\overline{t}+O (T^{d+2})
\ee
where $r_{max} (\boldsymbol\phi)$ is the value of the radial coordinate for which the base intersects the top at an angle $\boldsymbol\phi$, as shown in Figure \ref{fig:cone_plot}. To find this value we equate the time coordinates of the top and base surfaces to give 
\be
\frac{1}{2}f (\mathbf{x}_p,\boldsymbol\phi){r_{max}}^2 (\boldsymbol\phi)=-r_{max} (\boldsymbol\phi)+T
\ee 
We then solve this for $r_{max} (\boldsymbol\phi)$ and take the positive solution. The solution can be expanded in $T$ and is simply $r_{max}=T+O (T^2)$, with angular dependent terms contributing at $O (T^2)$. The $O (T^2)$ term will contribute at $O (T^{d+2})$ in the volume integral and so can be ignored. Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) allows us to evaluate the integral, wherein we find 
\begin{figure}[t]
  \centering
    {\includegraphics[scale=0.5,trim={1cm 5cm 1cm 6cm},clip]{coneplot3}}
     \caption{A $3$-dimensional representation of the region $\mathcal{X}_p$ in RNCs. The top part of $\partial\mathcal{X}_p$ can be approximated as a flat cone, and the base surface intersects the top part at a radial coordinate, $r_{max} (\boldsymbol\phi)$, which will in general be a function of the angles $\boldsymbol\phi$ (in $3$-dimensions there is one angle $\phi$). This function is found by projecting down from the intersection to the $\overline{t}$ plane. The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
     \label{fig:cone_plot}
\end{figure}
\be\label{eq:VolumeNoK}
V_\blacktriangle (p)
=\frac{V_{d-1}}{d}T^d\left (1-\frac{d}{2 (d+1)}\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}T\right)
+O (T^{d+2})
\ee
where the $\delta^{\ibar\jbar}$ comes from the fact that cross terms ($\ibar\neq \jbar$) vanish under the angular integration. The defining relations for $A^{i}_{\;\ibar}$ can now be rearranged to give $A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}=h^{ij} (p_0)$. In GNCs the extrinsic curvature on the surface is given by
\be\label{eq:K}
K
=g^{\mu\nu }\nabla_{\mu}n_{\nu}
=-\Gamma^{0}_{\;ij}h^{ij}=-\frac{1}{2}\frac{\dot{h}}{h}.
\ee
Substituting this into~\eqref{eq:VolumeNoK} we obtain 
\begin{align}
V_\blacktriangle (T,\mathbf x)
&=\frac{V_{d-1}}{d}T^d\left (1+\frac{d}{2 (d+1)}K (0,\mathbf{x})T\right)
+O (T^{d+2}) \label{eq:TopVolumeWithK}\\
V_\blacktriangledown (-T,\mathbf x)
&=\frac{V_{d-1}}{d}T^d\left (1-\frac{d}{2 (d+1)}K (0,\mathbf{x})T\right)
+O (T^{d+2}), \label{eq:BottomVolumeWithK}
\end{align}
having dropped the subscript $p$. Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\left\langle \BP{k}\right\rangle$ and $\left\langle \BF{k}\right\rangle$.
